---
title: "Data Science with R"
subtitle: "Lecture 06: Running R on AWS"
author: "Christine Choirat, cchoirat@gmail.com"
date: "March 28, 2016"
output: ioslides_presentation
---

## Objective

Analyze the airline on-time performance dataset on an Amazon instance

1. Sign-up for an account
2. Set up a machine
3. Install R, Rstudio, ...
4. Upload the data
5. Run a basic logit model

## Sign-up for a Free Tier account: https://aws.amazon.com/free/

- Many free products.

- Amazon EC2 "Amazon Elastic Compute Cloud"
    - "Resizable compute capacity in the Cloud."
    - 750 hours per month of Linux, RHEL, or SLES t2.micro instance usage
    - 750 hours per month of Windows t2.micro instance usage
    - For example, run 1 instance x 1 month or 2 instances x half a month
    - Expires 12 months after sign-up.
 
## Sign-up for a Free Tier account: https://aws.amazon.com/free/   

- Amazon S3
    - "Highly scalable, reliable, and low-latency data storage infrastructure."
    - 5 GB of Standard Storage
    - 20,000 Get Requests 
    - 2,000 Put Requests
    
- Amazon DynamoDB
    - "Fast and flexible NoSQL database with seamless scalability."

## Sign-up for a Free Tier account

## Sing-up for a Free Tier account

## SSH and key pairing

In a terminal:

```{r eval=FALSE}
chmod 600 Downloads/rstudio.pem
ssh -i Downloads/rstudio.pem ubuntu@ec2-54-175-200-1.compute-1.amazonaws.com
```

## Importing data

```{r eval=FALSE}
library("RStudioAMI")
linkDropbox()
excludeSyncDropbox("*")
includeSyncDropbox("Dropbox/DataAWS/")
```

## Running a model

```{r}
flights <- read.csv("Dropbox/DataAWS/allyears2k.csv")
names(flights)
glm(IsArrDelayed ~ Distance, data = flights, family = binomial("logit"))
```

## Amazon solutions

http://www.r-statistics.com/2015/06/setting-rstudio-server-using-amazon-web-services-aws-a-step-by-step-screenshots-tutorial/

https://aws.amazon.com/getting-started/launch-a-virtual-machine/

https://blogs.aws.amazon.com/bigdata/post/Tx3IJSB6BMHWZE5/Running-R-on-AWS

##  Array of system images (AMIs)

http://www.louisaslett.com/RStudio_AMI/

all instances support 64-bit

## Amazon EMR